---
title: "Plastics concetrations: Analysis-sediment"
output: html_notebook
---

```{r}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(tidyr)
library(reshape2)
```

Sediment: Artifical vs Nature pond
```{r}
plas <- read.csv("/Users/u0145079/Library/CloudStorage/OneDrive-KULeuven/Desktop/Doctorate/Publication/tables/MP_measurments/MPs_measurments_Sediment.csv")
```

Remove rows where Pond.Category is WWTP
```{r}
plas_filtered <- plas %>%
  filter(Pond.Category != "WWTP" & Location != "Sint-Donatus Park")
```

Calculating mean and standard error
```{r}
plas_filtered_averaged <- plas_filtered %>%
  group_by(Pond.Category) %>%
  summarise(
    mean = mean(Total_MPs_per_gram, na.rm = TRUE),
    sd = sd(Total_MPs_per_gram, na.rm = TRUE),
    count = n(),
    se = sd / sqrt(count) # Standard Error
  )
```

Plotting the bar chart with error bars

```{r}

plas_filtered_averaged$Pond.Category[plas_filtered_averaged$Pond.Category == "Artificial pond"] <- "City pond"
ggplot(plas_filtered_averaged, aes(x = Pond.Category, y = mean, fill=Pond.Category)) +
  scale_fill_manual(values = c("City pond" = "#cb5658", "Natural pond" = "#6885d0")) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "", y = "Mean MPs Number in a gram (MPs/gram) of Sediment", title = "Pond sediment samples") +
  theme_pubr() +
  theme(legend.position = "none")
```

```{r}
# Create a box plot with a dot overlay
ggplot(plas_filtered, aes(x = Pond.Category, y = Total_MPs_per_gram, fill = Pond.Category)) +
  geom_boxplot(outlier.shape = NA) + # Boxplot without outliers (dots will represent individual points)
  geom_jitter(color = "black", width = 0.2, size = 1.5, alpha = 0.6) + # Dot overlay (jitter)
  scale_fill_manual(values = c("City pond" = "#cb5658", "Natural pond" = "#6885d0")) +
  labs(x = "", y = "MPs per gram of Sediment", title = "Pond sediment samples") +
  theme_pubr() + # Use a minimal theme
  theme(legend.position = "none") # Remove the legend
```


Log transform the data

```{r}
plas_filtered$MPs_per_gram_log <- log(plas_filtered$Total_MPs_per_gram)
```

Check normality for each group after log transformation and removing NA values for Artifical pond
```{r}
plas_filtered$Pond.Category[plas_filtered$Pond.Category == "Artificial pond"] <- "City pond"
shapiro.test(na.omit(plas_filtered$MPs_per_gram_log[plas_filtered$Pond.Category == "City pond"]))
```

Check normality for each group after log transformation and removing NA values for Natural pond
```{r}
shapiro.test(na.omit(plas_filtered$MPs_per_gram_log[plas_filtered$Pond.Category == "Natural pond"]))
```

Perform two-sample t-test on log-transformed data
```{r}
t.test_result <- t.test(MPs_per_gram_log ~ Pond.Category, data = plas_filtered, var.equal = TRUE)
print(t.test_result)
```
Per pond analysis

```{r}
plas_filtered_averaged <- plas_filtered %>%
  group_by(Location) %>%
  summarise(
    mean = mean(Total_MPs_per_gram, na.rm = TRUE),
    sd = sd(Total_MPs_per_gram, na.rm = TRUE),
    count = n(),
    se = sd / sqrt(count) # Standard Error
  )

print(plas_filtered_averaged)
```

```{r}

plas_filtered_averaged$Category=c("Artificial pond","Natural pond","Natural pond","Natural pond","Natural pond","Artificial pond","Natural pond","Natural pond","Natural pond","Natural pond")

ggplot(plas_filtered_averaged, aes(x = Location, y = mean, fill = Category)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Artificial pond" = "#cb5658", "Natural pond" = "#6885d0")) +
  labs(x = "Pond", y = "Mean MPs number in a gram of sediment", title = "") +
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  coord_flip() # Swap the axes to make bars lay horizontally

```

Check for Homogeneity of Variances:
```{r}
library(car)
plas_filtered$Location <- as.factor(plas_filtered$Location)
leveneTest(MPs_per_gram_log ~ Location, data = plas_filtered, center=mean)
```
Check for normality
```{r}
shapiro.test(plas_filtered$MPs_per_gram)
```

Perform ANOVA
```{r}
anova_result <- aov(plas_filtered$MPs_per_gram_log ~ plas_filtered$Location)
summary(anova_result)
```

Post-hoc
```{r}
TukeyHSD(anova_result)
```

MICROPLASTICS COMPOSITIONS-SEDIMENT
```{r}
plas <- read.csv("/Users/u0145079/Library/CloudStorage/OneDrive-KULeuven/Desktop/Doctorate/Publication/tables/MP_measurments/Plastic-composition-Sediment.csv")

plas_norm = plas %>%
  dplyr::select(PE_PP_per_gram,PS_per_gram,PET_polyester_per_gram,Category,Location) %>%
  reshape2::melt()

plas_norm <- plas_norm %>%
  mutate(variable = dplyr::recode(variable, 
                           "PE_PP_per_gram" = "PE/PP",
                           "PS_per_gram" = "PS",
                           "PET_polyester_per_gram" = "PET/polyester"))

ggplot(plas_norm, aes(x = variable, y = value, fill = Category)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("High MPs" = "#cb5658", "Low MPs" = "#6885d0")) +
  labs(x = "Polymer type", y = "Mean MPs number in a gram of sediment", title = "") +
  theme_pubr() +
  theme(axis.text.x = element_text(vjust = 0.5, hjust=1),
        legend.position = "none") # This removes the legend completely

# Create the box plot without dot overlay
ggplot(plas_norm, aes(x = variable, y = value, fill = Category)) +
  geom_boxplot(outlier.shape = NA) +  # Box plot without outliers
  geom_jitter(color = "black", width = 0.2, size = 1.5, alpha = 0.6) + # Dot overlay (jitter)
  scale_y_log10() +  # Logarithmic y-axis to handle skewed data
  scale_fill_manual(values = c("High MPs" = "#cb5658", "Low MPs" = "#6885d0")) +  # Custom colors for categories
  labs(x = "Polymer type", y = "Log10 MPs number per gram of sediment", title = "") +
  theme_pubr() +
  theme(axis.text.x = element_text(vjust = 0.5, hjust=1),
        legend.position = "none")  # Removes the legend

# Create the box plot with dot overlay
ggplot(plas_norm, aes(x = variable, y = value, fill = Category)) +
  geom_boxplot(outlier.shape = NA) +  # Box plot without outliers
  geom_jitter(color = "black", width = 0.2, size = 1.5, alpha = 0.6) + # Dot overlay (jitter)
  scale_y_log10() +  # Logarithmic y-axis to handle skewed data
  scale_fill_manual(values = c("High MPs" = "#cb5658", "Low MPs" = "#6885d0")) +  # Custom colors for categories
  labs(x = "Polymer type", y = "Log10 MPs number per gram of sediment", title = "") +
  theme_pubr() +
  theme(axis.text.x = element_text(vjust = 0.5, hjust=1),
        legend.position = "none")  # Removes the legend
```

